Global Distribution of Unintended Pregnancies and Contraceptive Usage

### 2024-04-17 products/manuscript/Manuscript.qmd
Part 5 Updates
Added figure captions for both statistical analysis sections. Converted tibbles to organized tables for the rendered manuscript. Added a legend to the bootstrapping graph.

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. This process is similar to cross-validation because it relies on resamling the training data. Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We use a grid search to ensure that we try all combinations of the parameters in our grid. We complete the tuning process with a 5-fold cross-validation repeated 5 times. We added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so. Each of these processes are crucial in determining which model performs best for our dataset. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1.
```{r}
#| message: false
#| label: tbl-1
#| tbl-cap: "Table 1: Summary Table Of Numeric Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "Number of Countries With Percent Unintended Pregnancy Using SARC"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "Graph Depciting Percent Unintended Pregnancy by Standarized Count of Total Women"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "Graph Depicting Percent Unintended Pregnancies Using SARC by region"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "Violin Plot Depicitng Usage of SARc Per Region"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
```{r}
#| message: false
#| label: fig-5 - fig 10
#| fig-cap: "Predictor Variable Distributions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))

@@

```{r}
#| message: false
#| label: tbl-2
#| fig-cap: "Strongest Predictor: sarc_standard"
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @tbl-3.
```{r}
#| message: false
#| label: tbl-3
#| fig-cap: "Regional Predictor"
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 10 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

@@

We created one model with all 5 predictors using multiple linear regression, shown in @tbl-4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| message: false
#| label: tbl-4
#| fig-cap: "Multiple Linear Regression with 5 Predictors"
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
```
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @tbl-5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors performs the best with the lowest RMSE and MAE values. 
```{r}
#| message: false
#| label: tbl-5
#| fig-cap: "Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
```
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @tbl-6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| message: false
#| label: tbl-6
#| fig-cap: "Cross Validation Performance Metric Comparisons"
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 
```

Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance. As seen in @fig-11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line. The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line. While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data. 
```{r}
#| message: false
#| label: fig-11
#| fig-cap: "Evaluating Final Model Performance on Test Data"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```
We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data. The figure shown in @fig-12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model. The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction. For the predictions that do not fall on the expected 45 degree line, the confidence intervals do not intersect with the line either. These findings increase the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions. 
```{r}
#| message: false
#| label: fig-12
#| fig-cap: "Evaluation of Uncertainty in Full Model Predictions"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
```

Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. As seen in @tbl-7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model. The `sarc_standard` predictor still has the strongest relationship with the outcome. The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome. This process confirms that our final model should contain all 5 predictors. 
```{r}
#| message: false
#| label: tbl-7
#| fig-cap: "LASSO Regression Results"
#| echo: FALSE
readRDS(here("results","tables", "resulttable16.rds")) 
```
It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-13. As the regularization penalty increases past 1, we see that the RMSE value increases exponentially. The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1. When the regularization penalty increases to 10, the RMSE increases to almost 11. Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression based on an RMSE value slightly greater than 5, which i comparable to the RMSE found in the first regression model with all 5 predictors. 
```{r}
#| message: false
#| label: fig-13
#| fig-cap: "Parameter Tuning for LASSO Regression"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

After creating the random forest model, we found that the RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better. Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy. The plot of the predicted versus observed values shown in @fig-14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.  
```{r}
#| message: false
#| label: fig-14
#| fig-cap: "Predicted vs Observed Values in Random Forest Model"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "randomforestpred.png"))
```

Again, it is crucial to tune hyperparameters associate with machine learning models. Based on the red line displayed in @fig-15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors. This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model. However, there is a minimal increase in RMSE when all five models are included in the model with a node size of 1.  
```{r}
#| message: false
#| label: fig-15
#| fig-cap: "Tuning Parameters for Random Forest Model"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "randomforesttune.png"))
```

@@

Our full statistical analysis of the final model with 5 predictors included the standardized number of women using short acting reversible methods, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors. Cross-validation using LASSO regression and random forest modeling techniques did not result in an increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data. Based on the model using LASSO regression, the number of women using short acting reversible contraception is the strongest predictor of the outcome, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place. The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome. The percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors are all negatively associated with the outcome in this model. This implies that the percentage of unintended pregnancies decreases when a higher percentage of women are currently married and when the country experiences a higher maternal mortality rate based on abortion and unintended pregnancies. However, these effects are so small that it is likely these predictors do not have a real effect on the outcome. This analysis was successful in determining that region does have a significant effect on the percentage of unintended pregnancies due to short acting reversible methods with the Latin America/the Caribbean region experiencing the outcome at the highest rates.

>>>>

Step four is to critically analyze the performance of our models to determine which one performs the best in predicting our outcome, which is also done in the statistical-analysis.qmd file in the code > analysis-code folder. We chose 2 performance metrics to find for each of our 9 models: Root Mean Square Error (RMSE) and Mean Absolute Error (MAE). We chose these performance metrics because they are commonly used for linear regression models like the ones we have created. RMSE provides an average magnitude of the error in the predictors, and MAE provides an average absolute difference between the predicted and actual values. We include both metrics to increase the validity of our results. Once we decided on the final model, we use cross-validation to determine how well it will perform on data it has not seen before. This involves splitting the data into testing and training pieces , which allows us to compare the predictions created using both datasets. To examine the uncertainty in the model, we will use the bootstrapping method to resample our data 100 times to find confidence intervals for our predictions. This process is similar to cross-validation because it relies on resampling the training data. Lastly, we used a LASSO regression to add a penalty to each parameter included in the model to determine if any variables should be removed to improve model performance. It is important to tune the hyperparameters associated with LASSO regression to determine which model fits best. We use  a grid search to ensure that we try all combinations of the parameters in our grid. We complete the tuning process with a 5-fold cross-validation repeated 5 times. We added a random forest model to attempt to improve model performance. We used another parameter tuning grid to allow for cross-validated tuning of 2 variables: `mtry`, the number of variable to possibly split at in each node, and `min_n`, the minimal node size. We set the number of trees to 300 because it strikes a good balance between achieving the best performance metrics and the processing time required to do so. Each of these processes are crucial in determining which model performs best for our dataset. Result tables for each model and test are stored in the results > tables folder. Figures created from cross-validation, boostrapping, LASSO regression, and random forest modeling are stored in the results > figure > final-figures folder.

@@

After deciding on a single outcome variable, we created a summary table with that outcome and the seven predictors we will use to predict it, which can be seen in @tbl-1. [insert explanation here]
```{r}
#| message: false
#| label: tbl-1
#| tbl-cap: "Summary Table Of Numeric Variables"
#| echo: FALSE
predictortable<- readRDS(here("results", "tables", "predictortable.rds"))
predictortable
```

We created a multitude of exploratory plots to understand our data better. The first exploratory result shown in @fig-1 displays the distribution of the outcome variable: percent of unintended pregnancies due to short acting reversible methods. The `pct_upreg_sarc` variable is slightly skewed to the right with an average around 12%.
```{r}
#| message: false
#| label: fig-1
#| fig-cap: "[insert explanation here]"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_distribution.png"))
```

The next set of variables we explored were the total counts of women using each birth control method. There were two large outliers that prevented the graphs from being useful. After exploring the outliers, we standardized using the population size of reproductive aged women (15-49 years). There is a positive association between the percentage of unintended pregnancies among short acting reversible method users and the total number of women using that method of birth control, which is to be expected. Our outcome variable is the number of unintended pregnancies among women who are using short-acting reversible contraception, so the number of unintended pregnancies that can be classified in this outcome is dependent on the number of women using this type of contraception. This percentage should increase as the number of women using short acting-reversible contraception increases because more women are at risk of experiencing an unintended pregnancy due to this contraception if they are using this type of contraception in the first place. This outcome variable does not speak to the effectiveness of short-acting reversible contraception, so it makes sense that our outcome would increase as more women put themselves at risk of experiencing the outcome through becoming classified as short-acting reversible contraception users. As explained in the supplementary material, the dataset includes the percentage of unintended pregnancies among long-acting reversible contraception and sterilization users, traditional birth control users, and users who do not use a method of birth control. Each of these percentages are positively associated with number of people in its respective classification group that it represents. The association for the contraception that we chose to focus on for this analysis, short acting reversible contraception, is displayed in @fig-2.
```{r}
#| message: false
#| label: fig-2
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_totalsarc.png"))
```

We stratified these four graphs by region to examine how the relationship differs across each of the regions which is shown in @fig-3 . The graph for short acting reversible methods shows clearer differences between regions, and Latin America/the Caribbean leads in the percentage of unintended pregnancies in this category, but Africa shows the sharpest increase in percentage of unintended pregnancies as the total population increases.
```{r}
#| message: false
#| label: fig-3
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "pctupreg_sarc_stratified.png"))
```

We also created violin plots using the categorical usage variables for each birth control method as shown in @fig-4. It appears that the distribution of short acting reversible method usage is skewed towards lower percentages of unintended pregnancies across all usage levels, suggesting that more countries have lower percentage of unintended pregnancies among women using this method. The high usage countries appear to have a bottleneck in percentage of unintended pregnancies. The high and low usage countries appear to be skewed towards higher levels of unintended pregnancies, but the medium usage countries appear to have lower percentage of unintended pregnancies. 
```{r}
#| message: false
#| label: fig-4
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "sarc_pctupreg_violin.png"))
```

Lastly, we explored the predictor variables that will be used in our models, and the graphs are shown in @fig-5 - @fig-10. The Africa region accounts for most of the observations with about 53 countries, followed by Asia that includes about 35 countries. Latin America/the Caribbean accounts for about 25 countries. Europe and Oceania account for the least number of observations. The distribution of women currently married is approximately normal with an average around 57% and no major outliers. The distribution of women never married is slightly skewed to the right with an average around 30%, but there are a few outliers greater than 60%. The distribution of the total maternal mortality rate is highly skewed to the right with a couple outliers. We would have expected the highest rates to belong to India and China because they have the most people, but it is Chad, Sierra Leone, and South Sudan that account for the highest rates of total maternal mortality, which each belong to the African region. The distribution of the maternal mortality rate from unintended pregnancies is highly skewed to the right with a couple outliers, which is to be expected based on the trend shown in the previous histogram. Chad and South Sudan also account for the highest rates of maternal mortality in this category. The distribution of maternal mortality rate from abortions is also highly skewed to the right. The outliers are accounted for by Gambia and Mauritania in this predictor, which both belong to the African region.
```{r}
#| message: false
#| label: fig-5 - fig 10
#| fig-cap: "insert explanation here"
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "exploratory-figures", "region_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "currmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "nevmarried_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeaths_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsupreg_distribution.png"))
knitr::include_graphics(here("results", "figures", "exploratory-figures", "matdeathsabs_distribution.png"))

@@

```{r}
#| message: false
#| label: tbl-2
#| fig-cap: "This simple linear regresion model shows a strong effect of the standardized population size using short-acting reversible contraception methods on the outcome of percentage of unintended pregnancies due to short acting reversible contraception. This strong positive relationship indicates that as more women use this form of contraception, more women are susceptible to unintended pregancies due to SARCs because they are now at risk of experiencing the outcome. The coefficient of 87.65998 is the largest single coefficient of any predictor that we studied, which increases our confidence in including it as a predictor in our final model."
#| echo: FALSE
readRDS(here("results","tables", "resulttable1.rds")) 
```

The second model we created used the region predictor, which revealed that the Latin America/the Caribbean region had the strongest relationship with the outcome variable followed closely by the Asia region. The Oceania region does not appear to have a strong effect on the outcome variable with the Europe region only showing a slightly larger association. These associations can be seen in @tbl-3.
```{r}
#| message: false
#| label: tbl-3
#| fig-cap: "In a simple linear regression with the `region` predictor only, the Latin America/the Caribbean region has the strongest relationship with the outcome based on a coefficient of 1.6218844 followed closely by Asia. Europe has a smaller efffect on the outcome based on the coefficient of 0.3299391, but Oceania has the smallest effect on the outcome. This model chose the African region as the reference group, which means that it has the weakest association with the outcome compared to the other 4 regions."
#| echo: FALSE
readRDS(here("results","tables", "resulttable2.rds")) 
```
Details about the 10 different models we created can be found in the statistical-analysis.qmd file in the code > analysis-code folder along with an exploration of the correlation between predictors.

@@

We created one model with all 5 predictors using multiple linear regression, shown in @tbl-4. This model shows the strongest predictor is the number of women using short acting reversible contraception. The next strongest predictor is the European region followed by the Latin America/the Caribbean region and the Asia region, which is a shift from the model with the singular region predictor. The Oceania region now has a negative association with the outcome. The currently married predictor is associated with a weak decrease in the outcome. The coefficient estimates for the rate of maternal death due to unintended pregnancies and abortion both show a very weak positive association with the outcome. 
```{r}
#| message: false
#| label: tbl-4
#| fig-cap: "The multiple linear regression model with all 5 predictors confirms that the `sarc_standard` predictor has the strongest effect on the outcome, but the effect size has decreased slightly. The region variable is the second most impactful predictor, but there is a shift in which region has the strongest effect on the outcome. Africa still serves as the control region, but Europe has the largest effect size in this model followed by Latin America/the Caribbean. Oceania still has the weakest effect, and the effect is negative in this model compared to the positive effect seen in the simple linear regression model with the region predictor. The percent of currently married women, rate of maternal deaths due to uninteded pregnancies, and rate of maternal deaths due to abortion are very weak predictors with miniscule coefficients."
#| echo: FALSE
readRDS(here("results","tables", "resulttable8.rds")) 
To determine which model performed the best, we started by examining the RMSE and MAE performance metrics. We found that the model with the standardized total count of women using the short acting reversible contraception method performed the best in a single predictor model compared to the percent of currently married women and the rate of maternal mortality due to abortion. @tbl-5 shows the RMSE and MAE values for each of these three models compared to the null model with no predictors. The model with the population size using SARCs is the only single predictor model that performs better than the null. The full model with all 5 predictors performs the best with the lowest RMSE and MAE values. 
```{r}
#| message: false
#| label: tbl-5
#| fig-cap: "The performance metric comparisons clearly highlight the full model with 5 predictors as the best model with the lowest RMSE of 4.361363 and lowest MAE of 3.386193. The model with percent of currently married women and the model with the rate of maternal mortality due to abortion only perform slightly better than the null model, which just predicts the average outcome. These models are extremely poor. The best single predictor model is the one with the standard population size using SARCs with an RMSE of 5.143319 and an MAE of 3.953767."
#| echo: FALSE
readRDS(here("results","tables", "resulttable14.rds")) 
We used cross-validation of the two best models to determine how well the model performs on data it has not seen yet by splitting the data into testing and training pieces. The cross-validated results, shown in @tbl-6 lead to the same conclusion: the multiple linear regression model with 5 predictors performs better. The RMSE value for the single predictor model with `sarc_standard` is 4.822226, which is lower the the first RMSE value of 5.143319. While the RMSE value for the multiple linear regression model is slightly higher than the first time it was calculated, it is still lower than the single predictor model. Additionally, the standard error of the RMSE for the multiple linear regression model is 0.2768, which is lower than the standard error of the RMSE for the single predictor model of 0.3528. This means there is higher variance in the single predictor model. The R-squared value of 0.8206 for the multiple linear regression model is higher than 0.7861 for the single predictor model, which means the multiple linear regression model accounts for more of the variability in the data. Overall, the multiple linear regression model performs better.
```{r}
#| message: false
#| label: tbl-6
#| fig-cap: "Cross-validated (CV) performance metric comparisons were only completed on two models: the single predictor model with the `sarc_standard` variable and the full model with all five predictors because these two models performed the best in the first analysis. The RMSE metrics are closer using the CV technique with a differencec of 0.42 between the two models. The standard error of the full predictor model is slightly higher, but the R-squared value is also slightly higher for the full predictor model. These metrics confirm that the full predictor model is the best choice compared to single predictor models."
#| echo: FALSE
readRDS(here("results","tables", "resulttable15.rds")) 

Since we determined that the full model with all 5 predictors performs best, we graphed the predicted versus observed values using this model with both the testing and training data to further examine model performance. As seen in @fig-11, the training model predictions, shown in teal, hover pretty closely around the expected relationship between the outcome and the 5 predictors, which is depicted as the 45 degree line. The testing model predictions, shown in coral, are much more extreme than the training model predictions because they are farther from the 45 degree line. While there is some overlap between the two sets of predictions, it is likely that our model is overfitting the data. 
```{r}
#| message: false
#| label: fig-11
#| fig-cap: "Evaluating the final model performance on test data reveals that the model is strongly overfitting the data. The testing model predictions are shown in coral, and the variation in these observations is much higher compared to the training model predictions, shown in blue. Ideally, there would be complete overlap between the testing and training data predictions following the black dotted line that represent the ideal relationship between observed and predicted values. The majority of testing data predictions do not overlap with the training data prediction in this case, and the testing data predictions are father away from the preferred relationship shown by the black dotted line."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "fullmodelpred.png"))
```
We went one step further to examine the uncertainty in the full model predictions using the bootstrap method with 100 resamples of the training data. The figure shown in @fig-12 displays the 95% confidence intervals, shown as dark blue lines, for the predictions made by the model. The confidence intervals are narrow, which shows there is minimal uncertainty in the model prediction. For the predictions that do not fall on the expected 45 degree line, the confidence intervals do not intersect with the line either. These findings increase the validity of the conclusion that the model is overfitting the data because there is not much uncertainty in the predictions. 
```{r}
#| message: false
#| label: fig-12
#| fig-cap: "Evaluation of uncertainty in the full model predictions reveals decent overlap between the point estimates, shown in black triangles, and the median prediction observation, shown in coral circles. The error bars, shown in blue, represent 95% confidence intervals of the median estimates, and most of the confidence intervals contain the point estimates. This provides confidence in the certainty of the model's predictions, but this is not very meaninful considering how the full model overfits the data."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "bootstrappred.png"))
```

Since we have evidence that our model overfits the data, the last technique we want to use to explore our full model is Lasso regression. This technique could reveal predictors that should not be included in our final model, which could reduce the risk of overfitting. As seen in @tbl-7, the LASSO regression coefficients are much smaller than the first full model we created without the regularization penalty, but all 5 predictors remain in the model. The `sarc_standard` predictor still has the strongest relationship with the outcome. The percent of women currently married, rate of maternal deaths due to unintended pregnancies, and rate of maternal deaths to due to abortion all have weak negative associations with the outcome. This process confirms that our final model should contain all 5 predictors. 
```{r}
#| message: false
#| label: tbl-7
#| fig-cap: "LASSO regression results confirm the findings of the multiple linear regression model, but each of the coefficients has been shrunk as a result of the penalty parameter of 0.1. The `sarc_standard` predictor is still the most strongly related to the outcome, followed by the European region. The LASSO penaltty decreased the effect size of the Oceania region even further, and the smallest coefficient is now assigned to the percent of currently married women, which is a shift from the multiple linear regression model. The rates of maternal mortality due to unintended pregnancy and abortion are very weak predictors in this model as well."
#| echo: FALSE
readRDS(here("results","tables", "resulttable16.rds")) 
```
It is important to tune the hyperparameters associated with LASSO regression to determine the model that fits best, which is shown in @fig-13. As the regularization penalty increases past 1, we see that the RMSE value increases exponentially. The lowest RMSE for the model is slightly greater than 5 when the regularization penalty is less than 1. When the regularization penalty increases to 10, the RMSE increases to almost 11. Since we originally chose 0.1 as our penalty parameter, we found one of the best performing models with the LASSO regression based on an RMSE value slightly greater than 5, which i comparable to the RMSE found in the first regression model with all 5 predictors. 
```{r}
#| message: false
#| label: fig-13
#| fig-cap: "Parameter tuning for the LASSO regression reveals that the parameter should be smaller than 1 to ensure the lowest RMSE performance metric value. The RMSE value increases exponentially when the penalty parameter exceeds 1, so this provides confidence in our choice of using 0.1 as the penalty parameter in the prevoiusly displayed LASSO regression model. There is no evidence to use a smaller penalty parameter than 0.1 because it does not result in a lower performance metric."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "lassoparameters.png"))
```

After creating the random forest model, we found that the RMSE and MAE values are exactly the same because the same 5 predictors are used for the outcome, but the R-squared value shows that the random forest model performs slightly better. Since a higher R-squared value represents smaller differences between the observed and predicted values, we can conclude that this model is predicting our outcome with slightly higher accuracy. The plot of the predicted versus observed values shown in @fig-14 displays stronger correlation for predicted values compared to the testing and training done with cross-validation previously.  
```{r}
#| message: false
#| label: fig-14
#| fig-cap: "The graph of predicted versus observed values in the random forest model shows less deviation from the expected relationship visualized by the black dotted line in comparison to the multiple linear regression model displayed previously. While there are still several outliers as the predicted values increase past a value of 20, the observations hover around the expected relationship with conistency for values less than 10. While it is still likely that the random forest model is overfitting because it uses the same predictors, it appears that this model has higher predictive accuracy."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "randomforestpred.png"))
```

Again, it is crucial to tune hyperparameters associate with machine learning models. Based on the red line displayed in @fig-15 that shows the lowest possible RMSE score, our random forest model performs the best with a node size of 1 and 4 predictors. This begs the question of which one of the 5 predictors included in our model should be dropped to improve model performance because all previous performance metrics suggested that including all 5 predictors created the best model. However, there is a minimal increase in RMSE when all five predictors are included in the model, but the results suggest that a full model should have a node size of 4 to achieve the lowest RMSE.
```{r}
#| message: false
#| label: fig-15
#| fig-cap: "After tuning two parameters for the random forest model, it is clear that the best model based on the lowest RMSE performance metric value has a minimal node size of 1 for 4 out of the 5 models. This graph suggests that the final model should contain 4 predictors to achieve the best performance metric, but it does not provide information on which parameter should be dropped from the model. Interestingly, the best model with all 5 predictors has a minimal node size of 4."
#| echo: FALSE
knitr::include_graphics(here("results", "figures", "final-figures", "randomforesttune.png"))
```

@@

Our full statistical analysis of the final model with 5 predictors included the standardized number of women using short acting reversible methods, region, percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors. Cross-validation using LASSO regression and random forest modeling techniques did not result in an increase in performance metrics, and the model is strongly overfitting the data based on the predicted versus observed outcomes between the testing and training data. Based on the model using LASSO regression, the number of women using short acting reversible contraception is the strongest predictor of the outcome, which is to be expected because women must be using this form of contraception to have an unintended pregnancy due to failure of this contraceptive method in the first place. The most influential factor in this model is the region variable with European, Asian, and Latin American regions having the strongest effects on the outcome. This finding implies that these three regions have the best access to SARCs because they also have the highest standardized population size using this form of contraception in the first place. Since the African region was always selected as the control for the model, it is likely that this region has the lowest access to SARCs. The percent of currently married women, rate of maternal mortality due to abortion, and rate of maternal mortality due to unintended pregnancies as predictors are all negatively associated with the outcome in this model. This implies that the percentage of unintended pregnancies decreases when a higher percentage of women are currently married and when the country experiences a higher maternal mortality rate based on abortion and unintended pregnancies. However, these effects are so small that it is likely these predictors do not have a real effect on the outcome. This analysis was successful in determining that region does have a significant effect on the percentage of unintended pregnancies due to short acting reversible methods with the Latin America/the Caribbean region experiencing the outcome at the highest rates.
